---
title: "Homework 5 Branson Vu"
format: pdf
editor: visual
---

## Problem 1

```{r}
#| output: false
library(geoR)
library(rstan)
data("gambia")

# save mean of green 
green_mean <- mean(gambia$green)
```

## 1a.

```{r}
stan_data <- list(
  N = nrow(gambia),
  pos = gambia$pos,
  green = gambia$green,
  netuse = gambia$netuse,
  green_mean = green_mean
)
```

```{r}
# vectorizing the model instead of slide 9 approach so that my desktop/laptop can run the model faster

malaria_model <- "
data {
  int<lower=0> N;
  int<lower=0,upper=1> pos[N];
  vector[N] green;
  int<lower=0,upper=1> netuse[N];   
}
parameters {
  real beta0;
  real beta1;
  real beta2;
}
model {
  beta0 ~ normal(0, 10);
  beta1 ~ normal(0, 10);
  beta2 ~ normal(0, 10);

  pos ~ bernoulli_logit(beta0 + beta1 * green + beta2 * to_vector(netuse));
}
"
```

```{r}
#| output: false
malaria_fit <- stan(
  model_code = malaria_model,
  data = stan_data,
  chains = 2,
  iter = 6000,
  warmup = 1000,
  seed = 2727
)
```

```{r}
#extract posterior from model
posterior <- rstan::extract(malaria_fit)
```

## 1b

```{r}
log_odds0 <- posterior$beta0 + posterior$beta1 * green_mean + posterior$beta2 * 0
odds0 <- exp(log_odds0)

#summary
cat("Posterior mean log-odds (netuse=0):", mean(log_odds0), "\n")
cat("Posterior mean odds (netuse=0):", mean(odds0), "\n")

#histogram
hist(odds0,
     main = "Posterior Distribution of Odds (netuse=0)",
     xlab = "Odds of Malaria",
     col = "blue",
     border = "white",
     breaks =50
     )
```

## 1c

```{r}
log_odds1 <- posterior$beta0 + posterior$beta1 * green_mean + posterior$beta2 * 1
odds1 <- exp(log_odds1)

#summary
cat("Posterior mean log-odds (netuse=1):", mean(log_odds1), "\n")
cat("Posterior mean odds (netuse=1):", mean(odds1), "\n")

#histogram
hist(odds1,
     main = "Posterior Distribution of Odds (netuse=1)",
     xlab = "Odds of Malaria",
     col = "lightblue",
     border = "white",
     breaks =50
     )
```

## 1d

```{r}
#testing if bed-net decreases Malaria odds

probabilty_decrease <- mean(odds1 < odds0)
cat("P(bed-net decreases malaria odds)=", probabilty_decrease, "\n")

odds_ratio <- odds1 / odds0
median(odds_ratio)

quantile(odds_ratio, probs = c(0.025, 0.975))

```

The posterior probability that bed-net use decreases the odds of contracting malaria is 1 (100%), meaning every posterior sample supports this direction. The median odds ratio is approximately 0.5006761, suggesting that children who sleep under a bed-net have approximately half the odds of contracting malaria compared to children who do not, at average levels of greenness.

Since bed-net use (`netone`) is an observational variable and is not randomly assigned, confounding variables cannot be ruled out and causal conclusions should not be drawn. For example, families who use bed-nets may differ systematically from those who do not. Other unmeasured confounding variables include wealth, socioeconomic status, and healthcare access
